---
output: github_document
---



[![Travis-CI Build Status](https://travis-ci.org/rdinter/nassR.svg?branch=master)](https://travis-ci.org/rdinter/nassR)

An alternative for downloading various USDA data from [https://quickstats.nass.usda.gov/](https://quickstats.nass.usda.gov/) through R. You must sign up for an [API token](https://quickstats.nass.usda.gov/api) from the mentioned website in order for this package to work.

# Install

Development version (needs devtools installed):

```{r, eval = FALSE}
devtools::install_github("rdinter/nassR", upgrade = "never")
```

# Basic Usage

If a query works on the [https://quickstats.nass.usda.gov/](https://quickstats.nass.usda.gov/) interface, then it will work with the `nassR` package.

There are three basic functions for this package with the first as the workhorse:

1. `nass_data` this will return a data.frame to the specifications of the query from all of the arguments set in the function call. This mimics the simple "GET DATA" command off of [https://quickstats.nass.usda.gov/](https://quickstats.nass.usda.gov/) and requires an API token. There is a 50,000 limit for each call.
2. `nass_param` returns all of the possible values for a parameter in a query. Helpful to understand how to subset a query if it runs into the 50,000 limit.
3. `nass_count` returns the number of records for a query. Very useful in conjunction with `nass_param` to determine what queries can return data with a `nass_data` call.


## Key Install

```{r, include = FALSE, echo = FALSE}
library(nassR)
# set_nass_key("YOUR_KEY_IN_QUOTATIONS")
# First time, reload your enviornment so you can use the key without restarting R.
readRenviron("~/.Renviron")
# You can check it with:
Sys.getenv("NASS_KEY")
```

```{r, eval = FALSE, echo = TRUE}
library(nassR)
set_nass_key("YOUR_KEY_IN_QUOTATIONS")
# First time, reload your enviornment so you can use the key without restarting R.
readRenviron("~/.Renviron")
# You can check it with:
Sys.getenv("NASS_KEY")
```


The above script will add a line to your `.Renviron` file to be re-used whenever you are using the package. If you are not comfortable with that, you can add the following line to your `.Renviron` file manually to produce the same result.

`NASS_KEY='YOUR_KEY_IN_SINGLE_QUOTES'`

## Sample Code

### Example 1

Find the number of observations for county level variables in Wake County North Carolina.

```{r}
nass_count(state_name = "NORTH CAROLINA", county_name = "WAKE")
```

```
[1] 6997
```

### Example 2

View all of the parameter values for the group_desc category when the sector is CROPS.

```{r}
nass_param("group_desc", sector_desc = "CROPS")
```

```
[1] "CROP TOTALS"       "FIELD CROPS"       "FRUIT & TREE NUTS" "HORTICULTURE"      "VEGETABLES"       
```

### Example 3

Download the data pertaining to Wake County in 2012 for its value of agricultural land. Then look at the structure of the object returned by the function call.

```{r}
j5 <- nass_data(year = 2012, agg_level_desc = "COUNTY", short_desc = "AG LAND, INCL BUILDINGS - ASSET VALUE, MEASURED IN $", county_name = "WAKE")
str(j5)
```

```
'data.frame':	1 obs. of  39 variables:
 $ week_ending          : chr ""
 $ state_name           : chr "NORTH CAROLINA"
 $ country_code         : chr "9000"
 $ location_desc        : chr "NORTH CAROLINA, CENTRAL PIEDMONT, WAKE"
 $ begin_code           : chr "12"
 $ zip_5                : chr ""
 $ county_ansi          : chr "183"
 $ state_alpha          : chr "NC"
 $ util_practice_desc   : chr "ALL UTILIZATION PRACTICES"
 $ domain_desc          : chr "TOTAL"
 $ asd_desc             : chr "CENTRAL PIEDMONT"
 $ freq_desc            : chr "POINT IN TIME"
 $ prodn_practice_desc  : chr "ALL PRODUCTION PRACTICES"
 $ end_code             : chr "12"
 $ sector_desc          : chr "ECONOMICS"
 $ short_desc           : chr "AG LAND, INCL BUILDINGS - ASSET VALUE, MEASURED IN $"
 $ country_name         : chr "UNITED STATES"
 $ Value                : chr "806,464,000"
 $ reference_period_desc: chr "END OF DEC"
 $ CV (%)               : chr "7.9"
 $ class_desc           : chr "INCL BUILDINGS"
 $ asd_code             : chr "50"
 $ agg_level_desc       : chr "COUNTY"
 $ county_name          : chr "WAKE"
 $ region_desc          : chr ""
 $ watershed_desc       : chr ""
 $ state_ansi           : chr "37"
 $ congr_district_code  : chr ""
 $ domaincat_desc       : chr "NOT SPECIFIED"
 $ state_fips_code      : chr "37"
 $ group_desc           : chr "FARMS & LAND & ASSETS"
 $ watershed_code       : chr "00000000"
 $ unit_desc            : chr "$"
 $ source_desc          : chr "CENSUS"
 $ load_time            : chr "2012-12-31 00:00:00"
 $ county_code          : chr "183"
 $ statisticcat_desc    : chr "ASSET VALUE"
 $ commodity_desc       : chr "AG LAND"
 $ year                 : chr "2012"
```